HyperLogLog: Exponentially Bad in Adversarial Settings

### 2022-01-24 docs/HLL_exp_bad/main.tex
first set of small changes, referenced in red in reviews in google doc

We stress that none of our attacks are particularly deep in cryptanalytic terms. Yet HLL is being increasingly widely deployed, and never with protections against our attacks (insofar as we are aware). So despite their technical simplicity, we argue that our attacks are important in establishing the limitations of HLL in the broadest set of use cases. We present our attacks in an abstract way rather than tying them to any specific setting in order to illustrate their generality. We do give a more detailed exploration of their impact on the data-centre network monitoring system proposed in~\cite{flexswitch}, see Section~\ref{sec:exp}.

@@

\subsection{Paper Organisation}
The rest of the paper is organised as follows. Section~\ref{sec:related} immediately below discusses related work. Section~\ref{sec:overview} gives an overview of HLL. Section~\ref{sec:attacks} presents our different adversarial models. In Section~\ref{sec:manip}, we describe attacks on HLL and evaluate their impacts under the different adversarial models. Section~\ref{sec:redis} considers the security of the more modern version of HLL due to Ertl~\cite{hllnew} that is used in Redis. Our experimental results are discussed and related back to a specific application setting in Section~\ref{sec:exp}. Section~\ref{sec:formal} provides our formal security analysis of HLL. Section~\ref{sec:conclusions} provides our conclusions and ideas for future work.

@@

Both~\cite{cardestprivacy,hllvuln} propose mitigations using a salted HLL sketch, either as a replacement for (but at the cost of losing mergeability) or in addition to (inducing a memory overhead) an unsalted HLL. We show in Section~\ref{sec:formal} that when a salt is used appropriately (i.e.\ as the secret key for a pseudo-random function that is used in place of hash function $h$), security against adversarial inputs can be achieved even in our strongest adversarial setting. 
%In addition, our simulation-based security definition, in which the simulator in the ideal world has no access to the inserted values, automatically guarantees confidentiality of the values inserted into the HLL sketch, even in the strongest attack model considered in~\cite{cardestprivacy}. So, while our main objective is to protect against adversarial inputs, we also get privacy for free when using a keyed PRF in place of the HLL hash function.

Our work relates to the broader study of the performance of probabilistic data structures in adversarial environments.
Clayton \emph{et al.}~\cite{CCS:ClaPatShrPS19} recently provided a provable-security treatment of this topic. They used a game-based formalism to analyze Bloom filters, counting (Bloom) filters and count-min sketch data structures. They did not study HLL. Our use of a simulation-based security definition in place of the game-based ones used in~\cite{CCS:ClaPatShrPS19} is another point of distinction; see further discussion in Section~\ref{sec:formal}. Prior work also focused on Bloom filters~\cite{NaorY15,GerbetKL15} or flooding of hash tables~\cite{LiptonN93,CrosbyW03}. %References~\cite{GerbetKL15,CCS:ClaPatShrPS19} provide summaries of prior work in this area. 
Recent papers~\cite{Ben-EliezerJWY20,DBLP:journals/corr/abs-2011-07471,HassidimKMMS20,KaplanMNS21} have studied broad classes streaming algorithms under adversarial input from a theoretical viewpoint. Our applied viewpoint complements this work, focussing on the real-world usage of a specific cardinality estimator. 
%The authors of~\cite{Ben-EliezerJWY20} give a compiler that shows how to generically transform a non-robust streaming algorithm into a robust one in various scenarios. For the particular case of cardinality estimation, this results in solutions that are much less efficient than ours using a PRF; the security model used in this line of work is also weaker than our strongest model.

@@

\begin{itemize}
\item[S1:] The adversary does not know the details of the target HLL sketch but can access a shadow copy of the HLL sketch via its API. It can use the API to insert elements into this shadow HLL sketch, get its cardinality estimate, and reset it to its native empty state. The adversary does not know the state of the targeted HLL sketch, meaning in particular that it does not have the list of elements previously inserted into it by other users. It can insert items into the target HLL sketch, but is otherwise ``blind''. 
\item[S2:] The adversary has access to the details of the HLL implementation, i.e.\ it knows the number of buckets $m=2^n$ and the hash function $h$ in use. It can insert items into the HLL sketch, but is otherwise blind. In particular, an S2 adversary does not have access to the values of the HLL cardinality estimate at any point in its attack. This kind of attack scenario is appropriate in the context of the port scanning attack of~\cite{portscanhll}.
\item[S3:] The adversary has access to the details of the HLL implementation, i.e.\ it knows the number of buckets $m=2^n$ and the hash function $h$ in use. Additionally, it can access and interact with the sketch via an API provided by the sketch owner, allowing it to insert items into the HLL sketch and ask for the HLL's cardinality estimate at any point in time. However, it does not know the individual bucket contents.
\item[S4:] The adversary has direct access to the HLL sketch and all its internals, i.e., the number of buckets, the hash function $h$, and also the contents of each bucket in the HLL sketch \emph{at one specific point in time}, i.e.\ it is given a \emph{snapshot} of the sketch. It can access and interact with the sketch via an API provided by the sketch owner, as per scenario S3. 
%Note that, using these capabilities, an S4 adversary can compute a lower bound on the HLL cardinality estimate based on its initial snapshot and whatever it causes to be inserted into the HLL sketch during its attack (this is irrespective of whether its attack proceeds with input from other users being added to the HLL sketch in parallel with its attack or not; if it knows that there is no other input, then it can compute the \emph{exact} value of the HLL cardinality estimate at every point in its attack instead of a lower bound).
\end{itemize}

@@

However, this procedure does not work as simply as just described because of numerical issues that arise in the computation of the cardinality estimate. Specifically, increasing the value held in some bucket from $M[j]$ to $M[j]+1$ \emph{decreases} the denominator in the raw cardinality estimate $\sum_{j=1}^m2^{-M[j]}$ by an amount $2^{-M[j]-1}$. This does lead to a change in the raw cardinality estimate $E = \alpha_mm^2\cdot(\sum_{j=1}^m2^{-M[j]})^{-1}$. However, in practice, the raw cardinality estimate is rounded before being presented to the adversary, so this change may not actually be detectable, especially when $M[j]$ becomes large. We have done experiments which shows this to be the case for parameters arising in practice.

@@

For example, the rule could be to insert $x$ if and only if doing so does not increase the entry $M[j]$ stored in the relevant bucket beyond some pre-defined threshold value $v$. A simple computation based on the HLL raw cardinality estimator function in Fig.~\ref{fig:hll} shows that this rule would allow the insertion of a fraction of  $1-2^{-v}$ of the set $\mathcal{X}$ whilst keeping the raw cardinality estimate below $\alpha_m m 2^v$. This attack assumes $h$ behaves like a random function and exploits the geometric distribution of the position of the leading ``1'' bit in $w$, where $w = y_{n+1}y_{n+2}\ldots y_{n+1}$ and $y=h(x)$ (as in the HLL algorithm).

@@

\subsubsection{Initially Empty HLL Sketch, Scenarios S2 and S4:}

@@

As a second example, setting $B = 1$, we can expect to insert $|{\mathcal{X}}|/2^m$ items with a final cardinality estimate equal to $m \cdot \ln (1+\frac{1}{m-1})$, which is rounded to $1$ by the HLL algorithm (for $m$ of practical interest, e.g.\ $m=256$). Thus we keep the HLL cardinality estimate to a constant (1!) whilst being able to insert a constant fraction of the items from the input set ${\mathcal{X}}$.

\subsubsection{Initially Empty HLL Sketch, Scenario S1:}

@@

\paragraph{Core Idea:}
We assume the adversary again has (or is given) some set ${\mathcal{X}}$ of items from which to pick its inputs. The core idea of our attack is to use the first $B \leq B_{\max}$ items $x_1,\ldots, x_B$ from ${\mathcal{X}}$ to define a set of target buckets, and then to insert into the shadow HLL sketch further items $x \in {\mathcal{X}}$, checking for a cardinality estimate increase on each insertion. We initialise a list ${\mathcal{L}}$ containing $x_1,\ldots, x_B$. We add those items that do not increase the cardinality estimate on insertion into the shadow HLL sketch to ${\mathcal{L}}$ and reject the rest. We reset the shadow device when the cardinality estimate has increased too much, reinserting the first $B$ items from ${\mathcal{X}}$ again to re-establish a ``baseline'' of target buckets for the shadow HLL sketch. A key observation is that if in the attack we keep the number of ``active'' buckets to values $B^\prime$ that are always below $B_{\max}$, then the Hit Counting algorithm is used to compute the cardinality estimate; given its output $m \cdot \ln (m/V)$ (or, more exactly, a rounded version of this value) we can infer the exact value of $V$, and hence $B^\prime$. This enables precise detection of cardinality increases, provided we perform a reset before the Hit Counting algorithm stops being used (this is ensured by resetting when the estimate indicates that $B_{\max}$ buckets have been hit). After all of ${\mathcal{X}}$ has been processed as described, we take the constructed list ${\mathcal{L}}$ and reprocess it in the same way. We must ensure that the same set of initial items $x_1,\ldots, x_B$ is inserted into the shadow HLL sketch before any other items, so as to fix the same set of target buckets. The order of insertion can be otherwise randomised. % (and our analysis will take advantage of the fact that it can be). 
By repeated reprocessing, we will gradually remove items that do not hit the $B$ target buckets (defined by $x_1,\ldots, x_B$). Eventually, we obtain a final list ${\mathcal{L}}^*$ that we insert into the target HLL sketch.

@@

\paragraph{Simple Attack:}
The simple version of our attack uses the above idea, inserting items $x$ into the shadow HLL sketch but resetting it as soon as the cardinality estimate increases from $B$ to $B+1$ (indicating that a non-target bucket has been hit by item $x$). Acting in this way precisely determines whether each inserted item hits the target buckets or not, so has no ``false positives'' which need to be eliminated through reprocessing of the list ${\mathcal{L}}$. This means that the attack can be carried out in a streaming fashion in a single pass over ${\mathcal{X}}$. Assuming $h$ behaves like a random function, we must perform a reset with probability $\frac{m-B}{m} = 1 - \frac{B}{m}$ per item inserted, and hence in total an expected number of $(1 - \frac{B}{m})\cdot | \mathcal{X} |$ resets.

\paragraph{Complex Attack:}
The more complex version of the attack only resets the shadow HLL sketch when its cardinality estimate reaches a value indicating that $B_{\max}$ buckets have been hit (recall this number can be determined precisely because of the use of the Hit Counting algorithm for cardinality estimation). Hence, fewer resets should be required. However, when the number of active buckets has reached some value $B^\prime \leq B_{\max}$ this means that an item $x$ may not increase the cardinality estimate, yet still not hit one of the $B$ target buckets (instead it hits one of the other $B^\prime - B$ active buckets). Such an $x$ acts as a ``false positive" and needs to be eliminated through reprocessing of the list ${\mathcal{L}}$ (as previously described).

\paragraph{Analysis of the Complex Attack:}
%Suppose that after each reset, the order of the items in ${\mathcal{X}}$ (or ${\mathcal{L}}$) that have not yet been processed on the current pass is randomised. (Here we exclude the special items  $x_1,\ldots, x_B$ from the reordering.) Then 
On each pass, the probability that an item $x$ that should \emph{not} be inserted (i.e.\ $x$ is a true negative) does \emph{not} increase the cardinality estimate (i.e.\ $x$ is a false positive on this pass) is given by $(B^\prime - B)/m$, where $B^\prime$ is the current number of active buckets at the point when we try to insert $x$. (Here we ignore any inconvenient dependence issues across the passes.) This is bounded by $q := (B_{\max} - B)/m$ in the attack, since $B^\prime$ does not exceed $B_{\max}$. The probability that such an $x$ is correctly rejected is at least $1-q$. Hence the expected number of trial insertions done on $x$ (across different passes of the attack) before $x$ is correctly rejected is $1/(1-q) = m/(m - B_{\max} + B)$.

@@

\subsubsection{Initially Non-empty HLL Sketch, Scenario S2:}\label{sec:nonemptyHLLS2}
The adversary knows $h$ so it selects items $x$ from ${\mathcal{X}}$ for which the $\ell$ rightmost bits of $h(x)$ start with a ``1'' bit. In this case, since the relevant bucket $j$ already has $M[j] \ge 1$, the bucket's counter is not updated, and the cardinality estimate of the HLL sketch remains unchanged.

@@

\subsubsection{Initially Non-empty HLL Sketch, Scenario S4:}\label{sec:nonemptyHLLS4}
The adversary can adopt the same strategy as for S2, but relax the sampling requirements with the additional information available in this scenario (namely, the count values $M[j]$ per bucket in the HLL sketch). Specifically, it can allow some leading zero bits in the $\ell$ rightmost bits of $h(x)$ so long as this does not increase the current bucket estimate. In other words, if $h(x)$ is mapped to the $j$-th bucket containing $M[j] \ge 0$, then, instead of the strict requirement of having a ``1'' bit in the first position in the $\ell$ rightmost bits of $h(x)$, the adversary is satisfied when there is a ``1'' bit in any of the first $M[j]$ positions in this substring. (A value $M[j]=0$ then never permits insertions, meaning an empty bucket stays empty.) Under these conditions, the estimates in the buckets are never updated, and there is no effect on the cardinality estimate made by HLL.

@@

\subsubsection{Initially Non-empty HLL Sketch, Scenario S1:}

@@

It is natural to wonder whether a PRF is really required in place of hash function $h$ in order to make HLL secure, or whether a weaker cryptographic primitive would suffice. For example, since it has (roughly) uniform outputs, perhaps using a universal hash function (UHF) $h_K(\cdot)$ would be sufficient. In Appendix~\ref{sec:UHF} we show that, in general, UHFs do not suffice by providing a key recovery attack against HLL when $h$ is replaced by a specific UHF based on polynomial evaluation over a finite field.

@@

We want $F$ to be competitive with non-cryptographic hashing on small input sizes, as typical in current HLL implementations and applications (e.g.\ MurmurHash3 applied to an IP header). This makes off-the-shelf approaches like HMAC uncompetitive, since even for the shortest inputs, HMAC would require 4 evaluations of the underlying hash compression function.

@@

We have given a formal analysis of using a PRF as a countermeasure to our attacks. The use of a keyed function sacrifices HLL mergeability when the keys in the different HLLs are independently generated. It is an open problem to find methods for securing HLL sketches against adversarial inputs whilst retaining some form of mergeability. This may also require the introduction of entirely new data structures that trade increased storage for mergeability and improved security.

@@

\bibliographystyle{ieeetr}
\bibliography{ref.bib}

>>>>

We stress that none of our attacks are particularly deep in cryptanalytic terms. Fundamentally, they arise from implementations of HLL using a fixed hash function $h$ internally. Yet HLL is being increasingly widely deployed, and never with protections against our attacks (insofar as we are aware). So despite their technical simplicity, we argue that our attacks are important in establishing the limitations of HLL in the broadest set of use cases. We present our attacks in an abstract way rather than tying them to any specific setting in order to illustrate their generality. We do give a more detailed exploration of their impact on the data-centre network monitoring system proposed in~\cite{flexswitch}, see Section~\ref{sec:exp}.

@@

\subsection{Paper Organisation}
The rest of the paper is organised as follows. Section~\ref{sec:related} immediately below discusses related work. Section~\ref{sec:overview} gives an overview of HLL. Section~\ref{sec:attacks} informally presents our different adversarial models. In Section~\ref{sec:manip}, we describe attacks on HLL and evaluate their impacts under the different adversarial models. Section~\ref{sec:redis} considers the security of the more modern version of HLL due to Ertl~\cite{hllnew} that is used in Redis. Our experimental results are discussed and related back to a specific application setting in Section~\ref{sec:exp}. Section~\ref{sec:formal} provides our formal security model and security proof for HLL using a VIL-PRF. Section~\ref{sec:conclusions} provides our conclusions and ideas for future work.

@@

Both~\cite{cardestprivacy,hllvuln} propose mitigations using a salted HLL sketch, either as a replacement for (but at the cost of losing mergeability) or in addition to (inducing a memory overhead) an unsalted HLL. We show in Section~\ref{sec:formal} that when a salt is used appropriately (i.e.\ as the secret key for a pseudo-random function that is used in place of hash function $h$), security against adversarial inputs can be achieved even in our strongest adversarial setting. However, this comes at the cost of losing mergeability. 
%In addition, our simulation-based security definition, in which the simulator in the ideal world has no access to the inserted values, automatically guarantees confidentiality of the values inserted into the HLL sketch, even in the strongest attack model considered in~\cite{cardestprivacy}. So, while our main objective is to protect against adversarial inputs, we also get privacy for free when using a keyed PRF in place of the HLL hash function.

Our work relates to the broader study of the performance of probabilistic data structures in adversarial environments.
Clayton \emph{et al.}~\cite{CCS:ClaPatShrPS19} recently provided a provable-security treatment of this topic. They used a game-based formalism to analyze Bloom filters, counting (Bloom) filters and count-min sketch data structures. They did not study HLL. Our use of a simulation-based security definition in place of the game-based ones used in~\cite{CCS:ClaPatShrPS19} is another point of distinction; see further discussion in Section~\ref{sec:formal}. Prior work also focused on Bloom filters~\cite{NaorY15,GerbetKL15} or flooding of hash tables~\cite{LiptonN93,CrosbyW03}. %References~\cite{GerbetKL15,CCS:ClaPatShrPS19} provide summaries of prior work in this area. 
Recent papers~\cite{Ben-EliezerJWY20,DBLP:journals/corr/abs-2011-07471,HassidimKMMS20,KaplanMNS21} have studied broad classes streaming algorithms under adversarial input from a theoretical viewpoint. Our applied perspective complements this work, focussing on the real-world usage of a specific cardinality estimator. 
%The authors of~\cite{Ben-EliezerJWY20} give a compiler that shows how to generically transform a non-robust streaming algorithm into a robust one in various scenarios. For the particular case of cardinality estimation, this results in solutions that are much less efficient than ours using a PRF; the security model used in this line of work is also weaker than our strongest model.

@@

\begin{itemize}
\item[S1:] The adversary does not know the details of the target HLL sketch but can access a shadow copy of the HLL sketch via its API. It can use the API to insert elements into this shadow HLL sketch, get its cardinality estimate, and reset it to its native empty state. The adversary does not know the state of the targeted HLL sketch, meaning in particular that it does not have the list of elements previously inserted into it by other users. It can insert items into the target HLL sketch, but is otherwise ``blind''. 
\item[S2:] The adversary has access to the details of the HLL implementation, i.e.\ it knows the number of buckets $m=2^n$ and the hash function $h$ in use. It can insert items into the HLL sketch, but is otherwise blind. In particular, an S2 adversary does not have access to the values of the HLL cardinality estimate at any point in its attack. This kind of attack scenario is appropriate in the context of the port scanning attack of~\cite{portscanhll}, where the adversary has no API to the target HLL sketch but can cause entries to be inserted into it by a network monitor. 
\item[S3:] The adversary has access to the details of the HLL implementation, i.e.\ it knows the number of buckets $m=2^n$ and the hash function $h$ in use. Additionally, it can access and interact with the sketch via an API provided by the sketch owner, allowing it to insert items into the HLL sketch and ask for the HLL's cardinality estimate at any point in time. However, it does not know the individual bucket contents. This scenario corresponds to a client-server setting with the adversary playing the role of the client.
\item[S4:] The adversary has direct access to the HLL sketch and all its internals, i.e., the number of buckets, the hash function $h$, and also the contents of each bucket in the HLL sketch \emph{at one specific point in time}, i.e.\ it is given a \emph{snapshot} of the sketch. It can access and interact with the sketch via an API provided by the sketch owner, as per scenario S3. This scenario models settings where the service hosting the HLL sketch has been compromised by the adversary (and where the adversary continues to act as a client, as in S3).
%Note that, using these capabilities, an S4 adversary can compute a lower bound on the HLL cardinality estimate based on its initial snapshot and whatever it causes to be inserted into the HLL sketch during its attack (this is irrespective of whether its attack proceeds with input from other users being added to the HLL sketch in parallel with its attack or not; if it knows that there is no other input, then it can compute the \emph{exact} value of the HLL cardinality estimate at every point in its attack instead of a lower bound).
\end{itemize}

@@

However, this procedure does not work as simply as just described because of numerical issues that arise in the computation of the cardinality estimate. Specifically, increasing the value held in some bucket from $M[j]$ to $M[j]+1$ \emph{decreases} the denominator in the raw cardinality estimate $\sum_{j=1}^m2^{-M[j]}$ by an amount $2^{-M[j]-1}$. This does lead to a change in the raw cardinality estimate $E = \alpha_mm^2\cdot(\sum_{j=1}^m2^{-M[j]})^{-1}$. However, in practice, the raw cardinality estimate is rounded before being presented to the adversary, so this change may not actually be detectable, especially when $M[j]$ becomes large. We have done experiments which show this to be the case for parameters arising in practice.

@@

For example, the rule could be to insert $x$ if and only if doing so does not increase the entry $M[j]$ stored in the relevant bucket beyond some pre-defined threshold value $v$. A simple computation based on the HLL raw cardinality estimator function in Fig.~\ref{fig:hll} shows that this rule would allow the insertion of a fraction of  $1-2^{-v}$ of the set $\mathcal{X}$ whilst keeping the raw cardinality estimate below $\alpha_m m 2^v$. This attack assumes $h$ behaves like a random function and exploits the geometric distribution of the position of the leading ``1'' bit in $w$, where $w = y_{n+1}y_{n+2}\ldots y_{n+\ell}$ and $y=h(x)$ (as in the HLL algorithm).

@@

\subsubsection{Initially Empty HLL Sketch, Scenarios S2 and S4}

@@

As a second example, setting $B = 1$, we can expect to insert $|{\mathcal{X}}|/m$ items with a final cardinality estimate equal to $m \cdot \ln (1+\frac{1}{m-1})$, which is rounded to $1$ by the HLL algorithm (for $m$ of practical interest, e.g.\ $m=256$). Thus we keep the HLL cardinality estimate to a constant (1!) whilst being able to insert a constant fraction $1/m$ of the items from the input set ${\mathcal{X}}$.

\subsubsection{Initially Empty HLL Sketch, Scenario S1}

@@

\paragraph{\textbf{Core Idea}}
We assume the adversary again has (or is given) some set ${\mathcal{X}}$ of items from which to pick its inputs. The core idea of our attack is to use the first $B \leq B_{\max}$ items $x_1,\ldots, x_B$ from ${\mathcal{X}}$ to define a set of target buckets, and then to insert into the shadow HLL sketch further items $x \in {\mathcal{X}}$, checking for a cardinality estimate increase on each insertion. We initialise a list ${\mathcal{L}}$ containing $x_1,\ldots, x_B$. We add those items $x \in {\mathcal{X}}$ that do not increase the cardinality estimate on insertion into the shadow HLL sketch to ${\mathcal{L}}$ and reject the rest. We reset the shadow device when the cardinality estimate has increased too much, reinserting the first $B$ items from ${\mathcal{X}}$ again to re-establish a ``baseline'' of target buckets for the shadow HLL sketch. A key observation is that if in the attack we keep the number of ``active'' buckets to values $B^\prime$ that are always below $B_{\max}$, then the Hit Counting algorithm is used to compute the cardinality estimate; given its output $m \cdot \ln (m/V)$ (or, more exactly, a rounded version of this value) we can infer the exact value of $V$, and hence $B^\prime$. This enables precise detection of cardinality increases, provided we perform a reset before the Hit Counting algorithm stops being used (this is ensured by resetting when the estimate indicates that $B_{\max}$ buckets have been hit). After all of ${\mathcal{X}}$ has been processed as described, we take the constructed list ${\mathcal{L}}$ and reprocess it in the same way. We must ensure that the same set of initial items $x_1,\ldots, x_B$ is inserted into the shadow HLL sketch before any other items, so as to fix the same set of target buckets. The order of insertion can be otherwise randomised. % (and our analysis will take advantage of the fact that it can be). 
By repeated reprocessing, we will gradually remove items that do not hit the $B$ target buckets (defined by $x_1,\ldots, x_B$). Eventually, we obtain a final list ${\mathcal{L}}^*$ that we insert into the target HLL sketch.

@@

\paragraph{\textbf{Simple Attack}}
The simple version of our attack uses the above idea, inserting items $x$ into the shadow HLL sketch but resetting it as soon as the cardinality estimate increases from $B$ to $B+1$ (indicating that a non-target bucket has been hit by item $x$). Acting in this way precisely determines whether each inserted item hits the target buckets or not, so has no ``false positives'' which need to be eliminated through reprocessing of the list ${\mathcal{L}}$. This means that the attack can be carried out in a streaming fashion in a single pass over ${\mathcal{X}}$. Assuming $h$ behaves like a random function, we must perform a reset with probability $\frac{m-B}{m} = 1 - \frac{B}{m}$ per item inserted, and hence in total an expected number of $(1 - \frac{B}{m})\cdot | \mathcal{X} |$ resets.

\paragraph{\textbf{Complex Attack}}
The more complex version of the attack only resets the shadow HLL sketch when its cardinality estimate reaches a value indicating that $B_{\max}$ buckets have been hit (recall this number can be determined precisely because of the use of the Hit Counting algorithm for cardinality estimation). Hence, fewer resets should be required. However, when the number of active buckets has reached some value $B^\prime \leq B_{\max}$ this means that an item $x$ may not increase the cardinality estimate, yet still not hit one of the $B$ target buckets (instead it hits one of the other $B^\prime - B$ active buckets). Such an $x$ acts as a ``false positive" and needs to be eliminated through reprocessing of the list ${\mathcal{L}}$ (as previously described).

\paragraph{\textbf{Analysis of the Complex Attack}}
%Suppose that after each reset, the order of the items in ${\mathcal{X}}$ (or ${\mathcal{L}}$) that have not yet been processed on the current pass is randomised. (Here we exclude the special items  $x_1,\ldots, x_B$ from the reordering.) Then 
On each pass, the probability that an item $x$ that should \emph{not} be inserted (i.e.\ $x$ is a true negative) does \emph{not} increase the cardinality estimate (i.e.\ $x$ is a false positive on this pass) is given by $(B^\prime - B)/m$, where $B^\prime$ is the current number of active buckets at the point when we try to insert $x$. (Here we ignore any inconvenient dependence issues across the passes.) This is bounded by $q := (B_{\max} - B)/m$ in the attack, since $B^\prime$ does not exceed $B_{\max}$. The probability that such an $x$ is correctly rejected is at least $1-q$. Hence the expected number of trial insertions done on $x$ (across different passes of the attack) before $x$ is correctly rejected is $1/(1-q) = m/(m - B_{\max} + B)$.

@@

\subsubsection{Initially Non-empty HLL Sketch, Scenario S2}\label{sec:nonemptyHLLS2}
The adversary knows $h$ so it selects items $x$ from ${\mathcal{X}}$ for which the $\ell$ rightmost bits of $h(x)$ start with a ``1'' bit. In this case, since the relevant bucket $j$ already has $M[j] \ge 1$, the bucket's counter is not updated, and the cardinality estimate of the HLL sketch remains unchanged.

@@

\subsubsection{Initially Non-empty HLL Sketch, Scenario S4}\label{sec:nonemptyHLLS4}
The adversary can adopt the same strategy as for S2, but relax the sampling requirements with the additional information available in this scenario (namely, the count values $M[j]$ per bucket in the HLL sketch). Specifically, it can allow some leading zero bits in the $\ell$ rightmost bits of $h(x)$ so long as this does not increase the current bucket estimate. In other words, if $h(x)$ is mapped to the $j$-th bucket containing $M[j] \ge 0$, then, instead of the strict requirement of having a ``1'' bit in the first position in the $\ell$ rightmost bits of $h(x)$, the adversary is satisfied when there is a ``1'' bit in any of the first $M[j]$ positions in this substring. (A value $M[j]=0$ then never permits insertions, meaning an empty bucket stays empty.) Under these conditions, the estimates in the buckets are never updated, and there is no effect on the cardinality estimate made by HLL.

@@

\subsubsection{Initially Non-empty HLL Sketch, Scenario S1}

@@

It is natural to wonder whether a PRF is really required in place of hash function $h$ in order to make HLL secure, or whether a weaker cryptographic primitive would suffice. For example, since it has (roughly) uniform outputs, perhaps using a universal hash function (UHF) $h_K(\cdot)$ would be sufficient. In Appendix~\ref{sec:UHF} we provide a key recovery attack against HLL when $h$ is replaced by a \emph{specific} UHF based on polynomial evaluation over a finite field. This is enough to show that UHFs do not suffice in general.

@@

From inspection of Figure~\ref{fig:hll}, it is evident that hashing is the dominant cost when inserting an item into HLL. So we want $F$ to be competitive with non-cryptographic hashing, especially on small input sizes (as typical in current HLL implementations and applications, e.g.\ MurmurHash3 applied to an IP header). This makes off-the-shelf approaches like HMAC uncompetitive, since even for the shortest inputs, HMAC would require 4 evaluations of the underlying hash compression function.

@@

We have given a formal analysis of using a PRF as a countermeasure to our attacks. The use of a keyed function sacrifices HLL mergeability when the keys in the different HLLs are independently generated. It is an open problem to find methods for securing HLL sketches against adversarial inputs whilst retaining some form of mergeability. This may also require the introduction of entirely new data structures that trade increased storage for mergeability and improved security. It is also an open problem to extend our simulation-based approach to formal analysis to other data structures.

@@

\bibliographystyle{ieeetr}
\bibliography{./ref.bib}
